---
title: "Credit Card Fraud Detection"
output: pdf_document
---


# **Hypothesis Testing of Credit Card Fraud Detection**

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Credit card fraud detection is a mechanism of detecting the frauds such as stealing credit card information,  bankruptcy , credit card application & duplication fraud, identity theft to name a few.
There are several techniques to identify the defaulters who can possibly commit a credit card fraud. Some of them are:

1. Low Credit Score
2. Low annual income
3. Late credit payment difficulties
4. Occupation 
5. Credit history

For this purpose, let us consider the Kaggle dataset that consists of information regarding the customer details, their bank transactions , their occupation type , total annual income etc. 
In our analysis, we shall explore the low annual income as an attribute that describes the low credit history and failure to repay the installments of loan and credit amount borrowed from the bank to determine if an individual is a fraud or not.

## *Step 1 - Loading required Libraries*

```{r Declaring Packages,warning=FALSE,message=FALSE}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(tinytex)
library(readxl)
```

## *Step 2 - Reading the Data*

```{r Input the Data, echo=TRUE}
data <- read.csv("E://Classes Fall 2021//STAT//project//credit_card_dataset//application_data.csv")
data1 <- as_tibble(data)

```

## *Step 3 - Understanding the Data*

```{r Data summary and statistical information, echo=TRUE}
summary(data1)
```

Observation:
 
1. TARGET variable classifies the customers as  fraudster or not
TARGET = 1 tells that the customer is a fraudster
TARGET = 0 tells that the customer is not a fraudster
  
2. Let us pick the data fields that are important for our analysis and remove unwanted columns

## *Step 4 - Data Cleaning*

```{r Cleaning the data, echo=TRUE}
col_list <- c('SK_ID_CURR',
              'TARGET',
              'NAME_CONTRACT_TYPE',
              'CODE_GENDER',
              'FLAG_OWN_CAR',
              'FLAG_OWN_REALTY',
              'CNT_CHILDREN',
              'AMT_INCOME_TOTAL',
              'AMT_CREDIT',
              'AMT_ANNUITY',
              'NAME_TYPE_SUITE',
              'NAME_INCOME_TYPE',
              'NAME_EDUCATION_TYPE',
              'NAME_FAMILY_STATUS',
              'NAME_HOUSING_TYPE',
              'REGION_POPULATION_RELATIVE',
              'DAYS_BIRTH',
              'DAYS_EMPLOYED',
              'DAYS_REGISTRATION',
              'DAYS_ID_PUBLISH',
              'OWN_CAR_AGE',
              'FLAG_MOBIL',
              'FLAG_EMP_PHONE',
              'FLAG_WORK_PHONE',
              'FLAG_CONT_MOBILE',
              'FLAG_PHONE',
              'FLAG_EMAIL',
              'OCCUPATION_TYPE',
              'CNT_FAM_MEMBERS',
              'REGION_RATING_CLIENT',
              'REGION_RATING_CLIENT_W_CITY',
              'WEEKDAY_APPR_PROCESS_START',
              'HOUR_APPR_PROCESS_START',
              'REG_REGION_NOT_LIVE_REGION',
              'REG_REGION_NOT_WORK_REGION',
              'LIVE_REGION_NOT_WORK_REGION',
              'REG_CITY_NOT_LIVE_CITY',
              'REG_CITY_NOT_WORK_CITY',
              'LIVE_CITY_NOT_WORK_CITY',
              'ORGANIZATION_TYPE')

data_cleaned <- data1 %>% select(col_list)
data_cleaned <- data_cleaned %>% replace(is.na(.), 0)
data_cleaned <- data_cleaned %>% 
  filter(CODE_GENDER != "XNA")



```

1. Picked up selected columns
2. Replaced na with 0
3. Removed XNA from gender column

## *Step 5 - Exploratory Data Analysis *

### Summary statistics of Data

```{r Summary statisctics of Data, echo=TRUE}
str(data_cleaned)
summary(data_cleaned)
```

Mean Total Income = 168798

Median Total Income = 147150

```{r Percentage of total defaulters, ratio of female and male defaulters, echo=TRUE}
defaulters <- filter(data_cleaned,TARGET==1)
defaulters

percentage_defaulters <- nrow(defaulters)/nrow(data_cleaned)
percentage_defaulters*100

data_cleaned %>%
  group_by(CODE_GENDER) %>%
  summarise(fraction_defaulters = sum(TARGET)/n()*100)

data_cleaned['age_in_years'] <- floor(abs(data_cleaned$DAYS_BIRTH/365))


```
Observation:

8 percentage of the individuals are defaulters who have committed credit amount payment related frauds

## *Step 6 - Let us filter the dataset for defaulters or TARGET = 1*

```{r Defaulters, echo=TRUE}
defaulters <- filter(data_cleaned,TARGET==1)
defaulters

defaulters['age_in_years'] <- floor(abs(defaulters$DAYS_BIRTH/365))

```

## *Step 7 - Visualizations to understand the defaulters dataset*

```{r plot1, echo=TRUE}

ggplot(data = defaulters, aes(x=factor(TARGET))) +
  geom_bar(stat = "count", position = position_dodge()) +
  facet_grid(NAME_INCOME_TYPE ~ CODE_GENDER) +
  ggtitle("Total Number of Defaulters by their Income Type classified on Gender") +
  xlab("Defaulters") + ylab("Total Number of Defaulters count")


```
Observation:

1. Number of female defaulters are more than the male defaulters
2. Highest number of defaulters belong to the working professional category
2. Second highest defaulters are the pensioners, and comparitively the females are higher in this category compared to males

```{r plot2, echo=TRUE}

p <- ggplot(data=defaulters, aes(x=NAME_FAMILY_STATUS, fill=NAME_FAMILY_STATUS)) 
p + geom_bar(stat = "count") +
  ggtitle("Familial status of Defaulters - Working Professionals ") +
  xlab("Family Status") + ylab("Number of Defaulters Count ") + labs(fill = "Family Status") +
  coord_flip()


```
Observation:

The defaulters that are married have credit card payment difficulties than the single or unmarried individuals, for which the column HOUSING_TYPE tells that the working professionals have parents to look after which is one of the reasons why they have difficulty in payment.


```{r plot5, echo=TRUE}
defaulters[defaulters==""]<-NA
p <- ggplot(defaulters, aes(x=factor(1), fill=OCCUPATION_TYPE))
p + geom_bar() + coord_polar(theta="y") +
  ggtitle("Occupation status of Defaulters - Working Professionals ") +
  xlab(" ") + ylab(" ") + labs(fill = "Occupation Type")
```

Observation:

The majority and second highest of working professionals fall under Laborers that depend on  daily wages or weekly wages for their living,  

```{r plot3, echo=TRUE}
bwt_box <- ggplot(defaulters, aes(x=factor(TARGET),
                             y=age_in_years,color=factor(CODE_GENDER)))
bwt_box + geom_boxplot() +
  ggtitle(" Boxplots to compare the mean ages of Defaulters") +
       xlab("Defaulters") + ylab("Age in Years") 
```
Observation:

The mean age of female defaulters is above 40 years and more than the mean age of the male defaulters

```{r plot4, echo=TRUE}
scatter_plot <- ggplot(data=defaulters, aes(x=AMT_CREDIT, y=AMT_ANNUITY, color= factor(CODE_GENDER))) 
scatter_plot + geom_point() + 
  ggtitle("Scatter  plot  of  Credit Amount vs Loan Annuity ")+
  xlab("Credit Amount ") + ylab("Annuity Amount ")
```
Observation:

Even though there is huge credit amount that is borrowed by defaulters, they are unable to make regular payments in installments of the money towards their debts

## *Step 8 -  Hypothesis Testing*


There is a classification of Total Income threshold below which amount, the individuals are categorized as defaulters on the Kaggle discussions as claimed by one of the contributors to the project

Let us take random samples of different sizes to verify this claim

### Sampling process for Hypothesis Testing

```{r Randomnly pick samples of different sizes, echo=TRUE}

sample1 <-defaulters[sample(nrow(defaulters), 10), ]
sample2 <-defaulters[sample(nrow(defaulters), 25), ]
sample3 <-defaulters[sample(nrow(defaulters), 45), ]

```

### Let us check the distribution of Annual Income for the different samples

```{r Sampling Distribution, echo=TRUE}

sample_1_plot <- ggplot(sample1, aes(x=AMT_INCOME_TOTAL)) + 
  geom_density() +
  ggtitle("Sampling Distribution of Total Annual Income (n=10) ")+
  xlab("Credit Amount ") + ylab("Annuity Amount ")
sample_1_plot

sample_2_plot <- ggplot(sample2, aes(x=AMT_INCOME_TOTAL)) + 
  geom_density() +
  ggtitle("Sampling Distribution of Total Annual Income (n=25) ")+
  xlab("Credit Amount ") + ylab("Annuity Amount ")
sample_2_plot

sample_3_plot <- ggplot(sample3, aes(x=AMT_INCOME_TOTAL)) + 
  geom_density() +
  ggtitle("Sampling Distribution of Total Annual Income (n=45)")+
  xlab("Credit Amount ") + ylab("Annuity Amount ")
sample_3_plot

```

For large values of n, sampling distribution of Total Income appears to be normal
Here, for n= 45 the sampling distribution resembles a normal distribution

The mean Total Income for defaulters is 165611. 

## *Step 9 -  Formulating Null and Alternative Hypothesis*

Null hypothesis (H0) :- The null hypothesis is that population mean is 165611
Alternative Hypothesis (H1) :-Let us test alternative hypothesis to see if mean can be greater than 165611

## *Step 10 -  Test Statistic , calculating Z_observed*

```{r Test Statistic, echo=TRUE}
n <- nrow(sample2)

x_bar <- c()

for (i in 1:n){
  x_bar <- sum(sample2$AMT_INCOME_TOTAL)
}

x_bar <- x_bar/n

pop_sd <- sd(defaulters$AMT_INCOME_TOTAL)
pop_mean <- 165611

Zobs <- (x_bar-165611)/(746677/sqrt(45))
Zobs

```
Observation :

Z_observed value is less than the Z_alpha that is 1.96, that shows that Null hypothesis can be accepted

## *Step 11 -  Calculating p-value of Z_observed assuming alpha = 0.5 at 95% Confidence Level*

```{r p-vaue, echo=TRUE}
p_value <- pnorm(abs(Zobs))
p_value

```

## *Step 12 -  Statistical Conclusion*


Since p-value is greater than 0.05, accept the null hypothesis H0 and reject the alternative hypothesis H1

## *Step 13 - English Conclusion*

The mean annual income of the entire population is 165611 and the individuals will have an income less than this number to be classified as defaulters